---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Required libraries
```{r}
library(ggplot2)
```


## Loading and preprocessing the data
```{r}
d0 <- read.csv("activity.csv")

d1 <- na.omit(d0)

d1$date <- as.Date(d1$date)

# total number of steps for a day
totalSteps <- tapply(d1$steps, d1$date, sum)

```


## What is mean total number of steps taken per day?
### Histogram of the total number of steps taken each day
```{r}

hist(totalSteps, 
        main = "Distribution of steps per day", 
        xlab="Steps per day", 
        ylab="Frequency", 
        col = "red", 
        breaks = 25
    )

```


### Mean and median of the total number of steps taken per day
```{r}
# mean of total Steps
mean(totalSteps)
```

```{r}
# median of total Steps
median(totalSteps)
```


## What is the average daily activity pattern?

### Time seris plot of the interval and average number of steps taken in that interval 

```{r}

d2 <- aggregate(steps ~ interval, d1, mean)

par(mar = c(1.5, 1.5, 1, 1))

g1 <- ggplot(d2, aes(interval, steps)) + geom_line() +
        ggtitle("Average number of steps in a time interval") +
        xlab("Interval number") +
        ylab("Number of steps")

# plot the graph
g1

```

### Which 5 minute interval on average across all days containes the maximum number of steps
```{r}
# our data set already has the average number of steps for the interval
d2[which.max(d2$steps),]
```



## Imputing missing values
### Calculate and report the total number of missing values in the dataset
```{r}
sum(is.na(d0$steps))
```

### strategy for filling in all of the missing values in the dataset
#### We will use the average number of steps for the interval as replacement for missing value

```{r}
## function to return the replacement for the missing value
find.missing.val <- function(steps, interval){
                        calculated_steps <- NA
                        if (is.na(steps))
                            calculated_steps <- d2[d2$interval==interval, "steps"]
                        else
                            calculated_steps <- steps
                        return(calculated_steps)    
                    }

# Create a new dataset equal to the original but with the missing values filled in
d3 <- d0
d3$steps <-mapply(find.missing.val, d3$steps,d3$interval)


```

#### number of missing values in the new data set
```{r}
sum(is.na(d3$steps))
```



## Are there differences in activity patterns between weekdays and weekends?
```{r}

```
